---
title: "babynames"
author: "julia bloom"
date: "February 12, 2019"
output: html_document:
    keep_md: true
    self_contained: true
    toc: true
font-family: Helvetica
autosize: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)

library(babynames)
library(tidyverse)
library(ggplot2)
library(magrittr)
library(dplyr)
library(magrittr)
library(ggthemes)
```

# 1. Plot the most common names in 2017 over the entire period.

```{r, eval = TRUE}
top10_2017 <- babynames %>%
  filter(year==2017) %>%
  group_by(sex, name) %>%
  summarize(total = sum(n)) %>%
  arrange(desc(total)) %>%
  group_by(sex) %>%
  mutate(rank=row_number()) %>%
  filter(rank<=5)  %>%
  arrange(sex, rank)

top5f <- top10_2017 %>% filter(sex=="F")
top5m <- top10_2017 %>% filter(sex=="M")
```

The top 5 most common female names in 2017 were:

```{r}
top5f
```

The top 5 most common male names in 2017 were:

```{r}
top5m
```

Graph of the most common female and male names in 2017 across the entire period:

```{r, eval = TRUE}
babynames %>%
  filter(sex=='F') %>%
  filter(name %in% top5f$name) %>%
  ggplot(., aes(year, n)) +
  geom_line(aes(color=name, group=name)) +
  theme_tufte()

babynames %>%
  filter(sex=='M') %>%
  filter(name %in% top5m$name) %>%
  ggplot(., aes(year, n)) +
  geom_line(aes(color=name, group=name)) +
  theme_tufte()
```


# 2. Explore which names are most often used as unisex names. For which names has the popularity over time changed a lot?

```{r}
unisex <- babynames %>%
  group_by(year, name) %>%
  mutate(all_n = sum(n)) %>%
  ungroup() %>%
  mutate(unisex_prop = n / all_n)
```

```{r}
unisex$unisex <- ifelse(unisex$unisex_prop>0.2 & unisex$unisex_prop<0.8, 1, 0)
```

```{r}
top10_unisex <- unisex %>%
  filter(unisex == 1) %>%
  group_by(sex, name) %>%
  summarize(total = sum(n),
            unisex_prop = mean(unisex_prop)) %>%
  arrange(desc(total)) %>%
  group_by(sex) %>%
  mutate(rank=row_number()) %>%
  filter(rank<=5)  %>%
  arrange(sex, rank)
```

```{r}
top5f_unisex <- top10_unisex %>% filter(sex=="F")
top5m_unisex <- top10_unisex %>% filter(sex=="M")
```

The top five most common unisex names (among female babies) were:

```{r}
top5f_unisex
```

The top five most common unisex names (among male babies) were:

```{r}
top5m_unisex
```

Graphs of the most common unisex names among female and male babies across the entire period

```{r}
babynames %>%
  filter(sex=="F") %>%
  filter(name %in% top5f_unisex$name) %>%
  ggplot(., aes(year, n)) +
  geom_line(aes(color=name, group=name)) +
  theme_tufte()

babynames %>%
  filter(sex=="M") %>%
  filter(name %in% top5m_unisex$name) %>%
  ggplot(., aes(year, n)) +
  geom_line(aes(color=name, group=name)) +
  theme_tufte()
```

# 3. Identify one particular pattern in the data. Then try to capture this one pattern in a graphical display that highlights this one point.

I chose to explore whether notable presidents' first names saw a bump in popularity while they were in office. For the purposes of my analysis, I chose to look at the 5 most recent "notable" presidents: Theodore Roosevelt, Woodrow Wilson, Franklin D Roosevelt, Lyndon B. Johnson, and Barack Obama. I chose these presidents because I felt they had 

```{r, echo = TRUE}
babynames$presidential_names  <- ifelse((babynames$name=="Theodore") | (babynames$name=="Woodrow") | (babynames$name=="Franklin") | (babynames$name=="Lyndon") | (babynames$name=="Barack"), 1, 0)
```

```{r, echo = TRUE}
presidents <- babynames %>%
  filter(presidential_names=="1") %>%
  filter(sex=="M") %>%
  group_by(year, name)
```

The graph below shows the total number of male babies with the chosen presidential first names across the entire period.

```{r}
presidents %>%
  ggplot(., aes(year, n)) +
  geom_line(aes(color=name, group=name)) +
  theme_tufte()
```

Next, I decided to look at the percentage of male babies with presidential first names out of all male babies for each year.

```{r}
president_percent <- babynames %>%
  filter(sex=="M") %>%
  group_by(year) %>%
  mutate(total = sum(n)) %>%
  ungroup() %>%
  mutate(pres_prop = (n / total)*100)
```

The graph below shows the percentage of male babies with presidential names across the entire period.

```{r}
president_percent %>%
  filter(presidential_names=="1") %>%
  ggplot(., aes(year, pres_prop)) +
  geom_line(aes(color=name, group=name)) +
  theme_tufte()
```

