---
title: "QMSS-4015-Lab-1"
author: "Gregory M. Eirich"
date: "September 12, 2017"
output:
  html_document: default
---


## Lab 1
due Sept 27th
based on data from here:  http://www.thearda.com/Archive/Files/Codebooks/GSS2006_CB.asp

go line-by-line through this code; otherwise you might miss something

```{r set up}
setwd("/Users/gregoryeirich/Downloads")
gss = read.csv("GSS.2006.csv.xls")
```
```{r variable names}
names(gss) ## what variables are in here? "vpsu" and "vstrat" etc.
##View(gss) ## look at it as a spreadsheet
```



##1-  Use the table command, summary or describe commands, to tabulate one variable; and explain your results.

**How about I investigate something with work-life conflict?**

```{r table 1}
table(gss$jbintfam) ## tabulates "How often do you feel that the demands of your job interfere with your family life?" 
```
if I look up JBINTFAM at www.thearda.com above, I see that 1 = Always, and 5 = Never; there are 206 5's and only 51 1's

```{r table 2}
summary(gss$jbintfam) ## calculates this variable on a 1-5 scale; not as helpful
```
```{r table 3}
prop.table(table(gss$jbintfam)) ## this gives us the proportions for this variable

options(digits=2) ## what if I think those proportions are ugly?  Shorten them.

prop.table(table(gss$jbintfam)) ## this gives us the proportions for this variable
```
seeing this as a proportion helps a bit: ony 5% are Always feeling interfered with, while 20% are Never feeling interfered with

```{r table 4}
options(digits=7) ## what if I want to get back to my original level.  Default is 7.
```


```{r installation}
##install.packages("psych")  ## install this package, if you haven't already  
library(psych) ## this brings into your R session
```

```{r table 5}
describe(gss$wrkstat) ## this summarizes "Last week were you working full-time, part-time, going to school, keeping house, or what?" as if it were a number; this is not something you'd normally do
```

```{r table 6}
table(gss$wrkstat) ## Ahh, that's better: a table of categories: we see that 2322 people work full-time, for instance. 
```

``````{r table 7}
describe(gss$hrs1) ## this works better for a numeric variable like "IF WORKING, FULL OR PART TIME: How many hours did you work last week, at all jobs?".  The mean is 42.08.
```

## 2. Using either the plyr or doBy commands, compare the means and standard deviations for one variable, for at least two subgroups and explain your results. 

``````{r installation etc}
##install.packages("plyr")  ## install this package, if you haven't already
library(plyr)
```

``````{r tabulations 1}
ddply(gss, "jbintfam", summarise, Mean = mean(hrs1, na.rm = T), SD = sd(hrs1, na.rm = T))
```

ddply's results show that Always work 49.95 hours, while Nevers work 38.34 hours, with everyone else monotonically related.

``````{r installations 2}
##install.packages("doBy")  ## install this package, this is going to do the same thing as using plyr above
library(doBy)
```

``````{r tabulations 2}
summaryBy(hrs1~jbintfam, data=gss, FUN=c(mean, sd), na.rm=T) ## same result as above when using plyr; just another way
```

```{r boxplot, echo=TRUE}
boxplot(hrs1~jbintfam, data=gss) ## graphically displays the relationship in boxplot form
```

``````{r installations again}
##install.packages("ggplot2") ## this program will make a fancy boxplot instead; same results as above
library(ggplot2)
```

```{r ggboxplot, echo=TRUE}
p = ggplot(gss, aes(factor(jbintfam), hrs1))
p + geom_boxplot() + geom_jitter()
p + geom_boxplot(aes(fill = factor(jbintfam))) ## see how they include the NAs, which are the missings
```

##3. Using the gmodels or another command, crosstabulate two categorical or ordinal variables (getting proportions); and explain your results

``````{r installs}
##install.packages("gmodels")  ## install this package, if you haven't already
library(gmodels)
```

```{r final tabulation}
CrossTable(gss$jbintfam, gss$wrkstat, prop.r=F, prop.c=T, prop.t=F, prop.chisq=F, format="SPSS") ## this generates how frequently different types of workers feel interference between family and work
```

Interesting: full-timers experience 18% always conflicted, vs. part-timers experience 31% always conflicted.  Is that a statistically significant difference?  Stay tuned ...



